Moving Elements in List CRDTs

### 2020-04-06 paper/list-move.tex
Wording tweak
List move paper: camera-ready

\setcopyright{rightsretained}
\copyrightyear{2020}
\acmYear{2020}
\acmDOI{}
\acmConference[]{}{}{}
\acmBooktitle{}
\acmPrice{}
\acmISBN{}

@@

\begin{abstract}
Conflict-free Replicated Data Types (CRDTs) for lists allow multiple users to concurrently insert and delete elements in a shared list object.
However, existing algorithms behave poorly when users concurrently move list elements to a new position (i.e.\ reorder the list elements).
We demonstrate the need for such a move operation, and describe an algorithm that extends a list CRDT with an explicit move operation.
Our algorithm can be used in conjunction with any existing list CRDT algorithm.
In addition to moving a single list element, we also discuss the open problem of moving ranges of elements.
\end{abstract}

@@

All of the aforementioned algorithms allow replicas to \emph{insert} or \emph{delete} elements anywhere in the list.
However, none of them have explicit support for \emph{moving} elements from one position to another position in the list (\emph{reordering}).
This is a surprising omission because moving is a commonly required operation: in many to-do list applications, a user can drag and drop list elements to reorder them, and graphics software allows users to reorder the object list with commands such as ``bring to front'' (which moves an object to the top of the list, so that it occludes other objects) and ``send to back'' (which moves an object to the bottom, so that it is occluded by other objects).

@@

This short paper has made the case for list CRDTs to support an explicit move operation, allowing list elements to be reordered without duplicating them in the case of concurrent moves.
We have outlined an algorithm for single-element moves that can be retrofitted to any existing list CRDT that uses stable position IDs to refer to locations within the list (which is the case for all currently known list CRDTs).
We have also demonstrated the need for moving a contiguous range of list elements to a new position; finding an algorithm that adds this feature to a CRDTs is at present an open problem.
We hope that future research will succeed in developing an algorithm for moving ranges of elements.

>>>>

\copyrightyear{2020} 
\acmYear{2020} 
\setcopyright{acmlicensed}
\acmConference[PaPoC '20]{7th Workshop on Principles and Practice of Consistency for Distributed Data}{April 27, 2020}{Heraklion, Greece}
\acmBooktitle{7th Workshop on Principles and Practice of Consistency for Distributed Data (PaPoC '20), April 27, 2020, Heraklion, Greece}
\acmPrice{15.00}
\acmDOI{10.1145/3380787.3393677}
\acmISBN{978-1-4503-7524-5/20/04}

@@

\begin{abstract}
Conflict-free Replicated Data Types (CRDTs) for lists allow multiple users to concurrently insert and delete elements in a shared list object.
However, existing algorithms behave poorly when users concurrently move list elements to a new position (i.e.\ reorder the elements in the list).
We demonstrate the need for such a move operation, and describe an algorithm that extends a list CRDT with an explicit move operation.
Our algorithm can be used in conjunction with any existing list CRDT algorithm.
In addition to moving a single list element, we also discuss the open problem of moving ranges of elements.
\end{abstract}

@@

All of the aforementioned algorithms allow replicas to \emph{insert} or \emph{delete} elements anywhere in the list.
However, none of them have explicit support for \emph{moving} elements from one position to another position in the list (\emph{reordering}).
This is a surprising omission because moving is a commonly required operation: in many to-do list applications, a user can drag and drop list elements to reorder them, and graphics software allows users to reorder the object list with commands such as ``bring to front'' (which moves an object to the top of the list, so that it occludes other objects) and ``send to back'' (which moves an object to the bottom, so that it is occluded by other objects).
In the example of Figure~\ref{fig:smiley}, imagine the user first creates the pupils and then the white circles of the eyes; if the software by default places new objects in front of existing objects, the pupils would then not be visible.
The user thus needs to reorder the object list to bring the pupil objects to the front.

@@

This short paper has made the case for list CRDTs to support an explicit move operation, allowing list elements to be reordered without duplicating them in the case of concurrent moves.
We have outlined an algorithm for single-element moves that can be retrofitted to any existing list CRDT that uses stable position IDs to refer to locations within the list (which is the case for all currently known list CRDTs).
We have also demonstrated the need for moving a contiguous range of list elements to a new position; finding an algorithm that adds this feature to a CRDT is at present an open problem.
We hope that future research will succeed in developing an algorithm for moving ranges of elements.
